{\LARGE \textsc{Structural controlability of pollination networks}}

### 2016-03-02 paper/manuscript.Rnw
change structure of the paper files

\documentclass[a4paper,10pt]{article}
% abstract + title + authors
\usepackage{abstract}
\renewcommand{\abstractnamefont}{\normalfont\bfseries}
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} 
\usepackage{titlesec}
\usepackage{authblk}
\usepackage{datetime}
\newdateformat{usvardate}{
\monthname[\THEMONTH] \ordinal{DAY}, \THEYEAR}

@@

% graphics
\usepackage{graphicx}
\usepackage{float}

% references & links
\usepackage[
backend=biber,
style= authoryear,
citestyle = authoryear,
url=false,
doi=true,
isbn=false
]{biblatex}
\addbibresource{references.bib}

\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    citecolor=magenta,
    linkcolor=magenta
}
\usepackage{doi}
\begin{document}
\SweaveOpts{concordance=TRUE,prefix.string=figures/fig-}
\setkeys{Gin}{width=1\textwidth}

% title
% Article title
  \title{\fontsize{16pt}{10pt}\selectfont\textbf{Structural controlability of pollination networks}}

% Authors and Affiliations
  \author[1]{\large E. Fernando Cagua}
  \author[1]{\large Daniel B. Stouffer}
  \affil[1]{\normalsize Center of Integrative Ecology, School of Biological Sciences, University of Canterbury}

 \date{\small\usvardate\today}

@@

<<libraries and preparatory stuff, echo = FALSE>>=
library(ggplot2)
library(magrittr)

my_theme <- theme_classic() + theme(legend.position = "none", text = element_text(family = "Times"), axis.title.x = element_text(vjust = -0.7), axis.title.y = element_text(vjust = 1.5)) 
@

% Main Text
\section*{Introduction}

\todo[inline]{This intro was written when I though that the paper was mainly gonna be about the relative importance of species and how invasives fit there. Read with caution as I don't think that's the case anymore}

@@

Recent work on the control of complex systems suggest that in principle it is possible to alter any ecological community's composition, by modifying the abundances of just some key species. Here, we apply these theories to estimate the controllability of different ecological communities and to find the species, that due to the structural characteristics of their interactions are more likely to drive the dynamics of the community.

\section*{Methods \& Results}

Using six paired plant pollinator communities we explored the concepts of control theory. This is the initial data set things will be tested. If more data is considered necessary we'll see that later.

\subsection*{Controllability of ecological communities}

\subsubsection*{Minimum set of species }

Using maximum matching to control all species in the network. We found that depending on the scheme around of the species are needed to have a grasp at the whole community (\autoref{fig:n_species}). However, the relative ranking of networks remains relatively unchanged across schemes. So rather than focusing on the differences across schemes perhaps it would be interesting to see what characteristics make a network to require more or less driver nodes to be completely controled. Alternatively it would be interesting to see if there are systemic differences between different types of networks. An initial step would be to stay with bipartite networks of the mutualistic type (plant/frugivores) or the the consumer-resource type (host/paraiste or predator/prey). Perhaps some model networks of varying parameters?

\begin{figure}
\centering
<<fig = T, echo = F, width = 6, height = 3>>=
n_driver <- read.csv("../data/n_driver_species.dat")

ggplot(n_driver, aes(x = type, y = n_driver/n_species)) +
	geom_violin(colour = "grey") +
	geom_line(aes(colour = as.factor(net),group = as.factor(net))) +
	geom_point(aes(fill = as.factor(net)), shape = 21) +
	xlab("link structure scheme") + ylab("proportion of driver species") +
	my_theme
@
\caption{
\label{fig:n_species} 
Proportion of species that needs to be controled for all twelve networks under different link assumptions. AB indicates that the links go from plants to pollinators; BA that the links go from pollinator to plants; ``weight'' that the direction of control is chosen by dependency (ties are resolved by giving priority to plants in weight.A and to pollinators in weight.B); z-bi indicates that the links go in both directions}
\end{figure}

There was no significant difference between the number of species needed to control the system in an invaded and an uninvaded ecosystem.

When comparing the number of control nodes of the observed networks to their random counterparts an interesting pattern emegerges (\autoref{fig:random_n_species}). The number of control nodes is significantly larger than randomisations that mantain the degree distribution of the plants. However there is no significant difference when the randomisations mantain the degree of pollinators (or both plants and pollinators). This pattern is consistent across all schemes of link directionality. This suggest that the patterns of pollinator visitations are the structural characteristic that determines the controlability of the network. We probably should analyse other pollination networks to see if this pattern hold. We also should think about the implications of this finding.

\begin{figure}
\centering
<<fig = T, echo = F, width = 6, height = 3.5>>=
random_n <- read.csv("../data/random_n_driver.csv")
n_control <- n_driver %>%
	dplyr::select(-keep) %>%
	tidyr::separate(type, c("type", "keep"), "\\.", remove = T)

random_n_z <- random_n %>% 
	dplyr::group_by(net, type, keep, method) %>%
	dplyr::summarise(n_driver_u = mean(n_driver, na.rm = T),
									 n_driver_sd = sd(n_driver, na.rm = T)) %>%
	dplyr::full_join(n_control) %>% 
	dplyr::group_by() %>%
	dplyr::mutate(n_driver_z_score = (n_driver - n_driver_u)/n_driver_sd)

random_n_z %>% 
	ggplot(aes(x  = as.character(interaction(type, keep)), y = n_driver_z_score)) +
	geom_rect(ymin = -2, ymax = 2, xmin = -1, xmax = 10, fill = "grey90") +
	geom_hline(yintercept = 0, size = 0.4) +
	geom_vline(xintercept = seq(1.5, 5.5), linetype = 2) +
	geom_boxplot(aes(fill = method)) +
	scale_fill_discrete(name = "randomisations\nmantain\ndegree of", labels= c("plants", "pollinators", "both")) + 
	my_theme + 
	xlab("link structure scheme") +
	ylab("standard score") +
	theme(legend.position = "right",
				axis.line.x = element_blank(), axis.ticks.x = element_blank())
@
\caption{
\label{fig:random_n_species} 
Difference between the number of control nodes in the analysed networks versus randomisations of the same networks. Unshaded areas suggest a significant difference. Schemes same as in \autoref{fig:n_species}}
\end{figure}

It occured to me that it is possible that the patterns (and perhaps others observed like more phylogenetic signal in polinators than in plants) are a consequence of a greater diversity of pollinators in the interaction networks. More pollinators relative to plants imply that randomisations that mantain the degree distribution of pollinators are more constrained than those that only mantain the degree distribution of plants. It also means that statistical tests on that group have a more observations and therefore rendering significant results for smaller effect sizes. I cheked that with the 12 pollination networks of the study and there seems to be some evidence for that (\autoref{fig:zscore_ratio}). Perhaps it would be good to check with other networks to see if it holds.

\begin{figure}[h]
\centering
<<fig = T, echo = F, width = 6, height = 3.5>>=
net_info <- read.csv("../data/ntw_info.csv")

net_info %<>%
	dplyr::mutate(ratio_pol_pla = n_pol/n_pla)

random_n_z %>%
	dplyr::full_join(net_info) %>%
	dplyr::mutate(t = as.character(interaction(type, keep))) %>% 
	dplyr::filter(method == "c0", t != "BA.B") %>%
	ggplot(aes(x = ratio_pol_pla, y = n_driver_z_score)) +
	geom_point() +
	geom_smooth(method = "lm", alpha = 0.1, colour = "grey10") +
	xlab("ratio of number of pollinators to plants") +
	ylab("standard score") +
	my_theme + 
	facet_wrap(~t, scales = "free") + 
	theme(strip.background = element_blank())
@
\caption{
\label{fig:zscore_ratio} 
When considering only randomisations that mantain the degree distribution of plants, there seems to be a possitive relationship between the ratio of pollinators to plants and the Z-score of randomisations. The pattern is, however, not significant for the weighted schemes. The z-scores are identical for the AB.A and the BA.B schemes. Schemes same as in \autoref{fig:n_species}}
\end{figure}

\subsubsection*{Profiles of critical and redundant species}

Dividing the nodes and links among critical, ordinary and redundant we can have an idea of ``control profiles'' of the networks. Both species and interactions can be classified as redundant, ordinary and critical depending on the number of species that need to be controled to gain full control over the network once the species has been removed (\autoref{fig:s_crit_prop} and \ref{fig:i_crit_prop}).

\begin{figure}[tp]
\centering
<<fig = T, echo = F, width = 6, height = 2.5>>=
folder <- "../data/redundancy/node/"
node_redundancy <- list.files(folder, full.names = TRUE) %>%
    plyr::ldply(function(x, name){
        y <- read.csv(x) 
        info <- basename(x) %>% 
            stringr::str_sub(end = -5) %>% 
            stringr::str_split("_") %>% 
            unlist()
        y %>%
            dplyr::mutate(net = info[[1]],
                                        type = info[2],
                                        keep = info[3])
    })

nr_sum <- node_redundancy %>%
	dplyr::mutate(type = interaction(type, keep)) %>%
	dplyr::group_by(net, type, robustness) %>%
	dplyr::summarise(freq = n()) %>%
	dplyr::group_by(net, type) %>%
	dplyr::mutate(n_species = sum (freq)) %>%
	dplyr::group_by() %>%
	dplyr::mutate(prop = freq / n_species, 
								type = as.character(type), 
								robustness = as.character(robustness))

ggplot(nr_sum) +
	geom_boxplot(aes(x = type, y = prop, fill = robustness)) +
	geom_vline(xintercept = seq(1.5, 5.5), linetype = 2) +
	xlab("link structure scheme") +
	ylab("species proportion") +
	scale_fill_discrete(name = "nodes") +
	my_theme + 
	theme(legend.position = "right", 
				axis.line.x = element_blank(), 
				axis.ticks.x = element_blank())
@
\caption{\label{fig:s_crit_prop} Proportion of species that are ordinary (when removed the number of species needed to control the network doesn't change), redundant (when removed less species are needed) and critical (when removed the number of species needed to control the network increasses). Schemes same as in \autoref{fig:n_species}}
\end{figure}

\begin{figure}[tp]
\centering
<<fig = T, echo = F, width = 6, height = 2.5>>=
folder <- "../data/redundancy/edge/"
link_redundancy <- list.files(folder, full.names = TRUE) %>%
    plyr::ldply(function(x, name){
        y <- read.csv(x) 
        info <- basename(x) %>% 
            stringr::str_sub(end = -5) %>% 
            stringr::str_split("_") %>% 
            unlist()
        y %>%
            dplyr::mutate(net = info[[1]],
                                        type = info[2],
                                        keep = info[3])
    })

lr_sum <- link_redundancy %>%
	dplyr::mutate(type = interaction(type, keep)) %>%
	dplyr::group_by(net, type, robustness) %>%
	dplyr::summarise(freq = n()) %>%
	dplyr::group_by(net, type) %>%
	dplyr::mutate(n_species = sum (freq)) %>%
	dplyr::group_by() %>%
	dplyr::mutate(prop = freq / n_species, 
								type = as.character(type), 
								robustness = as.character(robustness))

ggplot(lr_sum) +
	geom_boxplot(aes(x = type, y = prop, fill = robustness)) +
	geom_vline(xintercept = seq(1.5, 5.5), linetype = 2) +
	scale_fill_discrete(name = "links") +
	xlab("link structure scheme") +
	ylab("interactions proportion") +
	my_theme + theme(legend.position = "right")
@
\caption{\label{fig:i_crit_prop} Proportion of interactions that are ordinary (when removed the number of species needed to control is reduced), redundant (when removed the number of species needed to control doesn't change) and critical (when removed the number of species needed to control the network increasses). Schemes same as in \autoref{fig:n_species}}
\end{figure}

On the species redundancy profile (\autoref{fig:s_crit_prop}), the profiles are also highly dependent on the chosen directionality scheme. For instance, in the case in which species have links in both directions there are not ordinary species and the number of critical species is relatively high, this is not surprising because under this scheme, there is a much smaller set of species that needs to be controled and therefore removing one of these can increase the minimum set of species needed to control the network.

However, there appears to be a relatively small variability between networks analysed using this approach. Again, as in the previous section, I think there is little that can be understood by using just this data-set. Again I think it would be most interesting to see if there are systemic differences between different types of networks where a larger variation in the profiles is expected.

On the interaction redundancy profile (\autoref{fig:i_crit_prop}) things seem a bit more interesting because the variation across networks and across link directionality schemes is pretty small. For instance for the bidirectional scheme, although the total number of links is doubled, the relative proportion of links that are critical and redundant is mantained.

\subsubsection*{Profiles of internal/external dilation}

\textcite{Ruths2014} proposed another clever way to characterise the control profile of a network by charactering the number of nodes that need to be controled among the ``mechanism'' that originates it. Similar as the maximum matching algorith, the idea comes from the concept that in the absense of cycles one node can at most control another one. Then the origin of control nodes can be splitted between nodes that need to be controled because they are a ``source node'' those that have only out links, control nodes associated to ``external dilations'' and those associated to ``internal dilations''.

It could be interesting to compare different networks using this framework. They argue that source dominated networks should be relatively easy to control in a top-down fashion. On the other hand internal dilation dominated networks represent the signature of a network that self-regulates and is not meant to be controled in that it's not designed to visit every possible state.

However as we can see here this framework is pretty useless when the direction of control is not perfectly known (as in most real wolrd networks really...). For instance when applying it to the data-set (\autoref{fig:dilation}), we see that when the control flows from pollinators to plants (AB.A) the network is source dominated simply because the pollinator diversity is greater than the plant's. When we asume that each guild can control each other (z-bi.all) the networks are internal dilation dominated because it's purely composed of cycles. However the weighted version might offer a meaningful possibility to develop the analysis.

\begin{figure}
\centering
<<fig = T, echo = F, width = 6, height = 3>>=
library(ggtern)
dilation <- read.csv(file = "../data/dilation_profiles.dat")

ggplot(dilation, aes(x = eta_s, z = eta_e, y = eta_i)) +
	geom_point(aes(fill = type.keep), shape = 21, size = 5, alpha = 0.7) +
	coord_tern() + theme_classic() + theme(text =  element_text(family = "Times"))

detach("package:ggtern", unload = TRUE, character.only = TRUE)
detach("package:ggplot2", unload = TRUE, character.only = TRUE)
library(ggplot2)
@
\caption{\label{fig:dilation} Control profiles of pollination networks. Control nodes are split among nodes that need to be controled because they are a source node, some that are control nodes because they respond to internal dilations (cycles) and some that are control nodes because they repsond to external dilations (sink nodes not matched)}
\end{figure}

\subsection*{Ranking species}

\subsubsection*{By frequency in control sets}

There are multiple sets of species with which is structurally possible to control network dynamics. An initial posibility os to relate the importance of a species for network dynamics to the frequency at which a species is present in the control sets.

Turns out unsurprisingly that the importance of particular species and its distribution over a networks dependends on the scheme chosen for the directionality of the links (\autoref{fig:n_control_dist}). 
When pollinators control plants (AB) no plant must be controled and most pollinators have an importance close to one. 
On the other hand when plants control pollinators (BA) presumably because the pollinator diversity is larger than the plant diversity, the pattern is less "bimodal" and intermediate levels of importance are allowed.

\begin{figure}
\centering
<<echo = F, cache = T>>=
folder <- "../data/maximum_matchings_summary/" 
files <- list.files(folder)
node_mm <- plyr::adply(files, 1, function(x){
	read.csv(file.path(folder, x)) %>%
		dplyr::mutate(net = unlist(stringr::str_split(x, "_"))[1],
									type = unlist(stringr::str_split(x, "_"))[2], 
									keep = unlist(stringr::str_split(x, "_"))[3],
									batch = unlist(stringr::str_split(x, "_"))[4])
}) %>%
	dplyr::select(-X1) %>%
	dplyr::group_by(net, type, keep, node) %>%
	dplyr::summarise(freq = sum(freq))
@

<<fig = T, echo = F, width = 6, height = 4>>=
ranked_mm <- node_mm %>%
	dplyr::filter(!is.na(freq)) %>%
	dplyr::group_by(net, type, keep) %>%
	dplyr::mutate(r = rank(freq, ties.method = "random"),
								f_f = freq/max(freq), 
								type.keep = as.character(interaction(type, keep)))  %>%
	dplyr::rename(net_name = net)

invasive_sp <- data.frame(node = c("p_4", "p_25"), invs.species = TRUE)

ranked_with_inv <- 
	ranked_mm %>% 
	dplyr::group_by() %>%
	dplyr::full_join(invasive_sp) %>% 
	dplyr::full_join(net_info)

ggplot(ranked_with_inv, aes(x = f_f, fill = net_name)) +
	geom_density(alpha = 0.025, adjust = 1, colour = "grey20") +
	geom_point(data = ranked_with_inv %>%
						 	dplyr::filter(invs.species), aes(y = scale(net) + 2),
						 shape = 21, fill = "grey") +
	facet_wrap(~type, ncol = 2) +
	my_theme +	
	xlab("species frequency in control sets") +
	theme(strip.background = element_blank(),
				axis.text.x = element_text(angle = 90))
@
\caption{\label{fig:n_control_dist} Distribution of the relative frequency a species is present in the set of control species among the networks. The frequency of invasive species in each of the six invaded networks is denoted with a circle. Results for the bidirectional liks (z-bi) were obatined only for the nine smallest networks, which included 3 invaded ecosystems}
\end{figure}

Among the weighted versions there was little difference if dependency ties were resolved for plants of for pollinators (\autoref{fig:n_control_dist_ntw}), again there are more species with intermediate levels of importance. Even more spread out are things when we consider bidirectional links. I imagine that this pattern is partly because there are many many more ways to control the network in the bidirectional scheme than in the others.

In both AB and BA the relative importance of different species is almost fully constrained by the scheme. In the weighted scheme, invasive species are in all cases among the most important ones. Interestingly the pattern is opposite for when the direction can go both ways. It would be good to see then whether these patterns are a reflection of the fact that invasive species are actually dominant or an artifact arising from the fact that all invasive species are plants and that all networks are dominated by pollinators. An initial inspection on the distributrion of plants and pollinators per network for these schemes suggest that that might be the case for the bidirectional scheme but not for the weighted one (\autoref{fig:n_control_dist_guild} and \ref{fig:n_control_dist_per_net}).

\begin{figure}
\centering
<<fig = T, echo = F, width = 6, height = 4>>=
ranked_with_inv %>%
	tidyr::separate(node, into = c("categ", "number"), sep = "_", remove = F) %>%
	ggplot(aes(x = f_f)) + 
	geom_density(aes(fill = categ), 
							 alpha = 0.5, adjust = 0.5, colour = "grey20") +
	geom_point(data = ranked_with_inv %>%
						 	dplyr::filter(invs.species), aes(y = scale(net) + 2), 
						 shape = 21, fill = "red") +
	facet_wrap(~type, ncol = 2) +
	scale_fill_grey(name = "", labels = c("pollinators", "plants")) +
	my_theme +	
	xlab("species frequency in control sets") +
	theme(strip.background = element_blank(),
				axis.text.x = element_text(angle = 90), legend.position = "right")
@
\caption{\label{fig:n_control_dist_guild} Distribution of the relative frequency a species is present in the set of control species by guild.}
\end{figure}

We explored that option by

<<>>=
ranked_with_inv  %>%
	tidyr::separate(node, into = c("categ", "number"), sep = "_", remove = F) %>%
	dplyr::group_by(net, type.keep, categ) %>%
	dplyr::summarise(f_f = median(f_f), 
									 ratio_pol_pla = mean(ratio_pol_pla)) %>%
	dplyr::group_by(net, type.keep) %>%
	dplyr::mutate(f_sum = f_f - lead(f_f)) %>%
	ggplot(aes(x = ratio_pol_pla, y = f_sum)) +
	geom_point() +
	geom_smooth(method = "lm")+ 
	facet_wrap(~type.keep)
@

\subsubsection*{By redundancy}

\subsubsection*{Strongly interacting species (interaction heterogeneity)}

\section*{Discussion}

For instance, restoration projects that aim to drive an invasive species out of the community, to reestablish a particular ecosystem service or in general modify the ecosystem state have had a limited amount of success. We argue that quantifying the relative dynamic importance of species the set of key species is the first step to evaluate current management efforts as well as to design targeted and informed

Dynamic system approach

\section*{Acknowledgmenets}

The authors thank Dr. Takeuki Uno for the insight provided to find the set of all maximum matching algorithms. EFC acknowledges the support from the University of Canterbury Doctoral Scholarship, the University of Canterbury Meadow Mushroooms Postgraduate Scholarship, a travel grant from the European Space Agency and a Rutherford Discovery Fellowship (to DBS). DBS ackloledges the support of a Rutherford Discovery Scholarship, administered by the Royal Society of New Zealand.

% \bibliography{references}
\printbibliography

\newpage
\section*{Supplementary Information}
\makeatletter 
\renewcommand{\thefigure}{S\@arabic\c@figure}
\makeatother
\setcounter{figure}{0}

>>>>

\documentclass[a4paper,10pt]{article}
\usepackage[margin=22.5mm]{geometry}

% abstract + title + authors
\usepackage{abstract}
\renewcommand{\abstractnamefont}{\normalfont\bfseries}
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} 
\usepackage{titlesec}
\usepackage{authblk}
\usepackage{datetime}
\newdateformat{usvardate}{
\monthname[\THEMONTH] \ordinal{DAY}, \THEYEAR}

@@

% graphics
\usepackage{graphicx}
\usepackage{float}
\usepackage{sidecap}
\sidecaptionvpos{figure}{t}
\usepackage[font=small]{caption}
\usepackage{subfig}
\usepackage{caption}

% table
\usepackage{booktabs}

% references & links
\usepackage[
backend=bibtex,
style= authoryear,
citestyle = authoryear,
url=false,
doi=true,
isbn=false
]{biblatex}
\addbibresource{references.bib}

\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    citecolor=magenta,
    linkcolor=magenta
}
\usepackage{doi}
\begin{document}

% title
% Article title
  \title{{\LARGE \textsc{Structural controlability of pollination networks}}}

% Authors and Affiliations
  \author{\large E. Fernando Cagua}
  \author{\large Kate Wooton}
  \author{\large Johanna Voinopol-Sassu}
  \author{\large Daniel B. Stouffer}
  \affil{\normalsize Center of Integrative Ecology, School of Biological Sciences, University of Canterbury}
 \date{}

@@

<<libraries, echo = FALSE, warning = F, message = F>>=
library(magrittr)
library(knitr)
library(ggplot2)
library(glmulti)
opts_knit$set (root.dir=normalizePath('../'))
@

<<auxiliary functions, echo = FALSE, warning = F, message = F>>=
"./code/V2.0/functions" %>% 
	list.files(full.names = T) %>%
	plyr::l_ply(source)

# figure counter
fig_count <- dplyr::data_frame(ref = "NA", count = 0)
fig_n <- function(ref) {
  fig_count <<- rbind(fig_count,
                      dplyr::data_frame(ref = ref,
                                        count = dplyr::last(fig_count$count) + 1))
  return(dplyr::last(fig_count$count))
}
fig_r <- function(ref) fig_count$count[fig_count$ref == ref]

# table counter
tbl_count <- dplyr::data_frame(ref = "NA", count = 0)
tbl_n <- function(ref) {
  tbl_count <<- rbind(tbl_count,
                      dplyr::data_frame(ref = ref,
                                        count = dplyr::last(tbl_count$count) + 1))
  return(dplyr::last(tbl_count$count))
}
tbl_r <- function(ref) tbl_count$count[tbl_count$ref == ref]
@

<<plot theme, echo = FALSE, warning = F, message = F>>=
fer_theme <- theme_bw() +
	theme(text = element_text(family = "Times"),
	      title = element_text(size = 7.5, hjust = 0),
	      legend.title = element_text(size = 9),
	      legend.text = element_text(size = 8),
				axis.text = element_text(size = 8),
				axis.title = element_text(size = 9, hjust = 0.5), 
				strip.text = element_text(size = 9, hjust = 0), 
				strip.background = element_blank(),
				plot.margin = grid::unit(c(5, 0, 2, 0), "mm"))
@

<<read data and results, echo = FALSE, warning = F, message = F>>=
net <- "./data/V2.0/networks" %>%
	read_networks()
meta <- readr::read_csv("data/ntw_info.csv") %>% dplyr::tbl_df() %>%
	dplyr::rename(count = method)
n_matched <- readRDS("./data/V2.0/n_matched.rds")
m_f <- readRDS("./data/V2.0/matching_frequency.rds")
r_w <- readRDS("./data/V2.0/random_weights.rds") %>% dplyr::tbl_df() %>%
  dplyr::rename_("net_name" = ".id")
r_d <- readRDS("./data/V2.0/random_directions.rds") %>% dplyr::tbl_df() %>%
  dplyr::rename_("net_name" = ".id")
s_p <- readRDS("./data/V2.0/species_properties.rds") %>% dplyr::tbl_df()
mod <- readRDS("./data/V2.0/species_models.rds")
@

\section*{Introduction}

@@

Recent work on the control of complex systems suggest that in principle it is possible to alter any ecological community's composition, by modifying the abundances of just some key species \autocite{Isbell2013, Cornelius2013}. Here, we apply these theories to estimate the controllability of different ecological communities and to find driver species: species, that due to the structural characteristics of their interactions are more likely to drive the dynamics of the community.

Invasive species have been shown to have a disproportionate effect on the structure of pollination communities. Influencing for example the strength of species interactions, and the degree of network nestedness and connectivity \autocite{Olesen2002, Aizen2008, Bartomeus2008, Vila2009, Traveset2013}. However whether this influence is translated into a driver role has not been tested. Here we use plant pollinator communities to investigate the number of species that should be managed to control population dynamics of the whole community, the characteristics that determine whether a species should be managed or not and how invasive species fit.

\section*{Methods}

To investigate the dynamic controllability of pollination networks, we used ten paired plants-pollinator communities. Each par was composed by a community invaded by a plant and a community free of the invasive species. Weighted visitation networks were constructed from previously published visitation data collected from pollination communities in Bristol, Great Britain \autocite{Lopezaraiza-Mikel2007} and the National Cap de Creus in Northeastern Spain \autocite{Bartomeus2008}. The four British uninvaded communities the non-invaded plots were obtained by experimentally removing the invasive species \textit{Impatients grandulifera}. Contrastingly, Spanish uninvaded communities were obtained from plots that were not been colonised by the invasive species \textit{Carpobrotus affine acinaciformis} or \textit{Opuntia stricta}. In each of the networks we calculated the minimum number of driver species---species that need to be managed in order to gain full control of the community.

We first assigned a direction to each link between plants and pollinators given by the direction of dependency. For each link between a plant and its pollinator we quantified the level of dependency of the plant on the pollinator and vice versa \autocite{Bascompte2006}. The link points to the plant if the its dependency on the pollinator is larger than the pollinator's dependency on the plant. The dependency of plant $i$ on pollinator $j$ is the ratio between the visits coming from pollinator $j$ and all pollinator visits to plant $i$. The dependency of pollinator $j$ on plant $i$ is the ratio of the visits by pollinator $j$ to plant $i$ and all visits of pollinator $j$.

In a directed network, a matching is a subset of links in which no two links share a common starting species or a common ending species. A species is matched if it is an ending node of an link in the matching. Otherwise, it is unmatched. It has been shown that the number of driver species can be calculated by counting the number of unmatched nodes in the directed graph representation of the pollination network \autocite{Liu2011}. We then found a maximum in the an alternative bipartite representation of the pollination network \autocite{Csardi2006}.

We calculated dependencies based on visitation frequency, which has been shown to be an appropriate surrogate of the inter-specific effects \autocite{Bascompte2006, Vazquez2005}. However, because the degree distribution, and ultimately the number of driver species, can be affected by the sampling method \autocite{Bluthgen2010} we compared the number of driver species in a pollination community in which both visitation frequency and effectiveness were measured \autocite{Ballantyne2015}.

To quantify to what extent the number of driver species is characteristic of the structure of pollination networks we compared them a suit of random null-models. A set of null models were based on network randomisations that maintained the degree of plants, pollinators, and both plants and pollinators. To analyse the effects of the chosen directionality, we devised a null model in which we randomised the visitation patterns and re-calculated the new dependencies. In all cases we computed 999 randomisations.

We quantified the relative importance of each species for network dynamics. To do this, we counted the number of times a particular species was a driver species across all possible maximum matchings in the pollination network. The number of maximum matchings were found by generating the line graph of the alternative bipartite network representation, and then enumerating the maximal cliques in the complement of the line graph \autocite{Csardi2006}.

Finally, we tackled the question whether some structural properties can predict the relative importance of driver species. Here, we regressed the importance of each species against measures of centrality (degree, betweenness), measures related to network robustness (contribution to nestedness) and measures of strength of association (visitation levels).

\section*{Results}

<<n_driver, echo = FALSE, warning = F, message = F>>=
# find n_driver
n_driver <- n_matched %>%
  dplyr::inner_join(meta) %>%
  dplyr::filter(!is.na(inv)) %>%
  dplyr::mutate(n_dr = n_pla + n_pol - n_matched,
                n_dr_p = 1 - n_matched / (n_pla + n_pol))

# Wilcoxon signed rank test
wt <- n_driver %>%
  dplyr::select(site, inv, n_dr_p) %>%
  dplyr::mutate(n_dr_p = qlogis(n_dr_p)) %>%
  tidyr::spread(inv, n_dr_p) %>%
  dplyr::select(-site) %>%
  as.matrix() %>% {
      wilcox.test(x = `[`(.,1), y = `[`(.,2), paired = T)
  }
@

We found that if we were to control the dynamics of the whole community, we would need to control between \Sexpr{round(min(n_driver$n_dr_p)*100)} and \Sexpr{round(max(n_driver$n_dr_p)*100)}\% (mean \Sexpr{round(min(n_driver$n_dr_p)*100)}\%) of the species in the community. The relative number of driver species did not change if the dependencies were assumed to be bidirectional, from plant to pollinators or from pollinators to plants (Supp. Information). However, the proportion of driver species did was not significantly different between invaded and uninvaded ecosystems (Wilcoxon signed-rank sum test, $p = \Sexpr{wt$p.value}$).

We examined the role that the asimmetry of the dependencies which defines the direction of control has a central role in structuring the controlability of polination networks. We found that empirical networks have a much larger  the importance of dependency

\begin{figure}
\centering
<<fig randomisations, echo = FALSE, warning = F, message = F, fig.height=2.2, fig.width=3.1>>=
r <- r_w %>%
  dplyr::bind_rows(r_d) %>%
  dplyr::inner_join(meta) %>%
  dplyr::filter(!is.na(inv)) %>%
  dplyr::mutate(n_dr = n_pla + n_pol - n_matched)

# non-parametric
# z_scores <- r$method %>%
#   unique() %>%
#   plyr::ldply(function(x){
#     n_driver %>%
#       dplyr::mutate(method = x)
#   }) %>%
#   dplyr::bind_rows(r) %>%
#   dplyr::group_by(net_name, method) %>%
#   dplyr::mutate(z_r = rank(n_dr, na.last = NA, ties.method = "average"),
#                 z_s = qlogis(z_r/n()),
#                 z_s = replace(z_s, is.infinite(z_s) & z_s > 0, qlogis(0.9999)),
#                 z_s = replace(z_s, is.infinite(z_s) & z_s < 0, qlogis(0.0001))) %>%
#   dplyr::filter(is.na(sim))

z_scores <- r %>%
  dplyr::group_by(net_name, method) %>%
  dplyr::summarise(mu = mean(n_dr, na.rm = T),
                   sigma = sd(n_dr, na.rm = T)) %>%
  dplyr::inner_join(n_driver) %>%
  dplyr::group_by() %>%
  dplyr::mutate(z_s = (n_dr - mu)/sigma)

z_scores %<>%
  plyr::ddply("method", function(x){
    x %>%
      dplyr::mutate(sign = t.test(z_s)$p.value)
  }) %>% 
  dplyr::group_by(method) %>%
  dplyr::mutate(m_z = median(z_s)) %>%
  dplyr::group_by() %>%
  dplyr::arrange(m_z)  %>%
  dplyr::filter(method %in% c("swsh_both", 
                              "quasiswap_count", 
                              # "r00_both", 
                              # "r0_both", 
                              # "c0_both", 
                              "r0"))

z_scores_method_order <- unique(z_scores$method)

z_scores %>%
  dplyr::mutate(method = factor(method, levels = z_scores_method_order),
                m_n = as.numeric(method),
                m_n = replace(m_n, m_n == 3, 3.5)) %>%
  ggplot(aes(x = m_n, y = z_s)) +
  # geom_hline(yintercept = c(-0), linetype = 2, alpha = 0.7, size = 0.25) + 
  geom_vline(xintercept = c(2.75), alpha = 0.7, size = 0.25) +
  geom_hline(yintercept = 0, size = 0.25, linetype = 2, colour = "grey30") +
  geom_boxplot(aes(group = method), size = 0.25, colour = "grey10",
							 outlier.shape = 21, outlier.size = 1) +
  geom_point(aes(y = 12, colour = sign <= 0.05), shape = 8, size = 1) +
  coord_flip() +
  scale_colour_manual(values = c(NA, "black")) +
  scale_x_continuous(breaks = c(1, 2, 3.5), labels = rev(c("dependency", "strength", "degree"))) + 
  xlab("") +
  ylab("number of driver species (z-score)") + 
  fer_theme +
  theme(axis.text.x = element_text(angle = 90), 
        legend.position = "none",
				plot.margin = grid::unit(c(2, 1, 2, 0), "mm"), 
				# panel.grid.minor.x = element_blank(),
				panel.grid.minor.y = element_blank())

@
\caption{\label{fig:randomisations} The (\textasteriskcentered indicates a significant difference)}
\end{figure}

<<importance, echo = FALSE, warning = F, message = F>>=
imp <- mod %>% 
  lapply(function(x){
    coef.glmulti(x, select = 0.95) %>%
      as.data.frame() %>%
      dplyr::add_rownames("factor") %>%
      dplyr::arrange(-Importance, factor)
  })

print_table <- . %>%
  dplyr::mutate(factor = paste0("$", factor, "$")) %>%
  magrittr::set_names(c("factor", "est.", "var", "No. mod.", "imp.", "C.I.")) %>% 
  dplyr::select(-var) %>%
  kable(format = "latex", align = "c", booktabs = T, digits = 2, escape = F, linesep = "") 
@

\begin{table}
\centering
\small
\subfloat[pollinators]{
<<table importance 1, echo = FALSE, warning = F, message = F, results = "asis">>=
print_table(imp[[1]])
@
}
\subfloat[plants]{
<<table importance 2, echo = FALSE, warning = F, message = F, results = "asis">>=
print_table(imp[[2]])
@
}
\caption{\label{table:sd} Minimum number and proportion of driver species necessary to control the full pollination network}
\end{table}

\begin{figure}
\centering
<<fig species, echo = FALSE, warning = F, message = F, fig.height=2.2, fig.width=3.1>>=
data_pol <- mod[[1]]@objects[[1]]$data
data_pla <- mod[[2]]@objects[[1]]$data

expand.grid(d = seq(min(data_pla$d), max(data_pla$d), by = 1),
            S_d = seq(min(data_pla$S_d), max(data_pla$S_d), length = 20), 
            n = mean(data_pla$n, na.rm = T), 
            S_v = mean(data_pla$S_v)) %>%
  dplyr::mutate(q = predict.glmulti(mod[[2]], 
                                    select = 0.95, 
                                    newdata = .)$averages %>%
                  as.vector(), 
                p = plogis(q)) %>%
  ggplot(aes(x = d, y = S_d, z = p)) +
  geom_tile(aes(alpha = p), fill = "#0e324b") +
  # scale_fill_gradient(name = "relative\nimportance", high = "#f0f0f0", low = "#525252") +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  xlab("degree") +
  ylab("strength of dependencies") +
  stat_contour(colour = "black", size = 0.75) +
  fer_theme + theme(panel.grid = element_blank())
@
\caption{\label{fig:prediction} aSS }
\end{figure}

\printbibliography

\end{document}
